\MakeUppercase{Notes on Layered $H$-Partitions of $k$-Planar Graphs} \thanks{This work was partly funded by NSERC and MRI.}

### 2019-05-03 k-planar.tex
Seems to work

We claim that $K[x]$ is rooted at $x$, i.e., $x$ is the lowest common ancestor of all nodes in $B_x$.  To see why this is so, suppose $K[x]$ has root $a\neq x$.  This implies that $a$ has two children $b\neq c$ where $x$ is a descendant of $b$, some node $y$ is a descendant of $c$, and there is an edge $vw\in E(G)$ such that $v\in S_x\subset V(N_x)\subseteq V(N_b)\setminus V(F_a)$
and $w\in S_y\subset V(N_y)\subseteq V(N_b)\setminus V(F_a)$.  Therefore, $v$ and $w$ are in the interior of two different faces of the Mercedes graph $M_a$.  Therefore the curve $vw$ intersects $M_a$.  If it intersects $M_a$ at some vertex in $Y_a$, then $v$ would be in $S_a$, not $S_x$, a contradiction.  If it intersects $M_a$ at some vertex in $F_a$, then $v$ would be in $S_{a'}$ for some ancestor $a'$ of $a$, also a contradiction.

@@

Now, consider the cubic graph $A$ homeomorphic to the union of $F_{a'}$ for all ancestors $a'$ of $x$.  It is helpful to think of this graph as follows:  We begin with $F_{a_0}$ where $a_0$ is the root of $K$.  We then cut $F_{a_0}$ into two pieces $F^{\bar{x}}_{a_0}$ and $F^x_{a_0}$ where $F^x_{a_0}$ contains $F_x$.  This process continues by partitioning $F_{a_1}:=F^x_{a_0}$ recursively.  The process terminates after some number $r$ of iterations, when $F^x_{a_r}=F_x$.

>>>>

We claim that $K[x]$ is rooted at $x$, i.e., $x$ is the lowest common ancestor of all nodes in $B_x$.  To see why this is so, suppose $K[x]$ has root $a\neq x$.  In short: This happens because some $v\in V(N_x)$ is incident on an edge $vw\in E(G)$ that crosses $Y_a$.  Longer: This implies that $a$ has two children $b\neq c$ where $x$ is a descendant of $b$, some node $y$ is a descendant of $c$, and there is an edge $vw\in E(G)$ such that $v\in V(N_x)\setminus V(F_x)\subseteq V(N_b)\setminus V(F_a)$
and $w\in S_y\subset V(N_y)\subseteq V(N_b)\setminus V(F_a)$.  (Note, we allow $x=b$ and/or $y=c$ here.)  Therefore, $v$ and $w$ are in the interior of two different faces of the Mercedes graph $M_a$.  Therefore the edge $vw$ intersects $M_a$.  If it intersects $M_a$ at some vertex in $Y_a$, then $v$ would be in $S_a$, not $S_x$, a contradiction.  If it intersects $M_a$ at some vertex in $F_a$, then $v$ would be in $S_{a'}$ for some ancestor $a'$ of $a$, also a contradiction.

@@

Now, consider the cubic graph $A$ homeomorphic to the union of $F_{a'}$ for all ancestors $a'$ of $x$. See \figref{A}.  It is helpful to think of this graph as follows:  We begin with $F_{a_0}$ where $a_0$ is the root of $K$.  We then cut $F_{a_0}$ into two pieces $F^{\bar{x}}_{a_0}$ and $F^x_{a_0}$ where $F^x_{a_0}$ contains $F_x$.  This process continues by partitioning $F_{a_1}:=F^x_{a_0}$ recursively.  The process terminates after some number $r$ of iterations, when $F^x_{a_r}=F_x$.

\begin{figure}
  \begin{center}
    \includegraphics{figs/A}
  \end{center}
  \caption{The graph $A$.}
  \figlabel{A}
\end{figure}
